\TheTitle

### 2017-07-19 paper/model.tex
Fixing the "commit sizes" sentence
correcting the algorithm
Improves the description, provides an example, a new figure, and
corrects the algorithm.

Notice that the DAG loses information about the master branch and the
repository that the master branch is part of. The \mt view of
this DAG is visible in Figure~\ref{fig:repoTree}. Note that the
direction of the edges of the DAG have been inverted, instead of
pointing from the child to the parent, it points from the parent to its
children, forming a path to the master branch. Also note that the DAG
has been simplified, showing only a single edge on the path to master
for any commit.

@@

        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
  \end{tikzpicture}
  \caption{Example of a sequence of events performed in different
    repositories. The horizontal axis represents time. Each horizontal
    section represents a different branch and/or repository. Each commit
    points to its parent. The intial commit is at time $t_0$, and the
    head is at $t_8$.}
  \label{fig:repoEvents}
%\vspace{-3mm}
\end{figure}

@@

@@

  \end{tikzpicture}
  \caption{\mt view of the commits  represented in
    Figure~\ref{fig:repoEvents} showing the path they followed to reach
    the master branch. In this model the successors of each commit
    represents the path followed by that commit to reach the master
    branch.}
  \label{fig:repoTree}
\vspace{-3mm}
\end{figure}

@@

Computing the \mt from a DAG for any repository may not be
possible; however, certain features of the development process of Linux
make it feasible to compute the \mt for the Linux repository.
First, the master branch of Linux is maintained by Linus Torvalds, and
only Linus has write access to it. We have verified this assertion in
previous research~\cite{German2015}. We have developed a heuristic that
is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
identifies the commits made directly to the master branch, whereafter it
recursively determines the shortest path (in terms of time), using the
DAG, from each commit to the master branch using the inverted DAG.

\begin{algorithm}
        \caption{Computing the \mt of Linux from the DAG}\label{fig:alg}
        \begin{algorithmic}
                \Function{ComputeMergeTree}{DAG}: tree
                \State {\# Compute the tree from the DAG of Linux repository.}
                \State {\# Returns $Tree$, a graph containing every commit }
                \State {\# in DAG with the path it followed to master.}

                \State $head \gets \textit{Head of master of git repository}$
                \State $master \gets \textit{traverse DAG from head using }$
                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
                \State $nodes(Tree) \gets nodes(DAG)$
                \State \Function{distance2Master}{cid} : seconds
                \State {\# Helper function}
                \State {\# Recursively compute shortest distance to master}
                \State {\# setting cid's successor (next) in its way to master.}
                \State {\# This function should be memoized. Otherwise it}
                \State {\# would run in exponential time.}
                \If {\textit{cid in master}}
                \State \Return 0
                \EndIf
                \State    $d \gets 	\infty$
                \State {\# Traverse the inverted DAG}
                \For{$c \in children(cid, DAG)$}
                \If {$c \in master$}
                \State $d_1 \gets commitTime(c)-commitTime(cid)$
                \Else
                \State {$d_1 \gets distance2Master(c)$}
                \EndIf
                \If {$d_1 < d $}
                \State $next \gets c$
                \State  $d \gets d_1$
                \EndIf
                \EndFor
                \State {\# $c$ is the commit that follows $cid$}
                \State {\# in its way to master}
                \State add edge $(cid, next)$ to $Tree$
                \State \Return $d$
                \EndFunction

                \State {\# Compute the distance for each commit}
                \State {\# discarding result}
                \For{$c \in nodes(DAG)$}
                \State $distance2Master(c)$
                \EndFor
                \State \Return $Tree$
                \EndFunction
        \end{algorithmic}
\end{algorithm}

@@

\begin{figure}[htbp]
        \centering
        {\fontsize{7}{9}
        \begin{verbatim}
Merge: 8cbd84f fd8aa2c
Author: Linus Torvalds <torvalds@linux-foundation.org>
Date:   Tue Aug 10 15:38:19 2010 -0700

@@

* 'for-linus' of git://neil.brown.name/md: (24 commits)
md: clean up do_md_stop
[... edited for the sake of space]
md: split out md_rdev_init
md: be more careful setting MD_CHANGE_CLEAN
md/raid5: ensure we create a unique name for kmem_cache...
...
        \end{verbatim}}\vspace{-5mm}
        \caption{Example of how merges record a subset of commits being merged. The
                commit only shows the first 20 one-line summaries messages for the 24
                non-merge commits it merged. The ending ``\ldots'' is part of the log
                and represents that other commits were merged.}
        \label{fig:sampleMerge}
\end{figure}

We used this information to evaluate the accuracy of the \mt model
extracted from the DAG\@. The method we followed started with the
extraction of the commit history up to July 20, 2016 form the Linux
repository. We computed the \mtof every commit until that date.
Since Linus Torvalds mostly does merging directly into the master
branch, we assumed that every merge by Linus is the root of a \mt, later
detecting merges that do not merge into the master branch and removing
them from the set of root merges. As described above, the log of a
merge-commit usually contains the number of commits in the merge the
first 20 summaries of commits being merged. We extracted merges by Linus
Torvalds using the command \verb|git log --merges --author='Torvalds'|
and compared the number of commits according to the log with the number
of commits in the \mt rooted in this commit. We also used the summaries
of the commits found in the merge (not necessarily all---see above) to
make sure those commits were in their corresponding \mt For
example, for the merge in Figure~\ref{fig:sampleMerge} we would expect
that the \mt rooted at \mycode{8cbd84f} contains 24 commits, and
the one-line summaries corresponds to commits in that \mt. We also
inspected those with differences to make sure they were true errors. The
results can be summarized as follows:

>>>>

The DAG does not retain information about where a commit was originally
created. The \mt form (Figure~\ref{fig:repoTree}) does not completely
rebuild the lost information, but is able to re-create the sequence of
merges required to merge the event, and the depth, in the tree, of that
event. The \mt does not include information about how the repository
events that are the children of a merge are related. The order of the
children does not matter.

Combining the DAG and the \mt results in a hybrid structure
(Figure~\ref{fig:repoDAGTree}) that no only shows which merges a commit
passes through, it rebuilds the order of those commits. In the example
shown, this hybrid structure makes it clear that merge 9 merges the
changes made in 6 and 8 into the commit 7. Finally, 11 merges merge 9
into the master branch.

@@

        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
  \end{tikzpicture}
  \caption{Example of a sequence of events performed in different
    repositories. The horizontal axis represents time. The branches and
    repositories are aligned horizontally, and color-coded. Each commit
    points to its parent. The initial commit is at time $t_0$, and the
    head is at $t_8$.}
  \label{fig:repoEvents}
%\vspace{-3mm}
\end{figure}

@@

\begin{figure}[htpb]
  \begin{center}
    \begin{tikzpicture}[auto, on grid, semithick, node distance=1cm, state/.style={circle, text=black, minimum size=7mm}]

      \node[state, draw=chartblue] (1) {1};
      \node[state, draw=chartblue, right=of 1] (4) {4};
      \node[state, draw=chartblue, right=of 4] (11) {11};
      \node[state, draw=chartblue, right=2.5cm of 11] (12) {12};

      \node[state, draw=chartyellow, above right= 1cm and 0.5cm of 11] (7) {7};
      \node[state, draw=chartyellow, above left= 1cm and 0.5cm of 11] (5) {5};
      \node[state, draw=chartyellow, right=of 7] (9) {9};
      \node[state, draw=chartyellow, left=of 5] (2) {2};
      \node[state, draw=chartmagenta, above=of 5] (3) {3};
      \node[state, draw=chartmagenta, above left=1cm and 0.5cm of 9] (6) {6};
      \node[state, draw=chartmagenta, above right=1cm and 0.5cm of 9] (8){8};
      \node[state, draw=chartyellow, above=of 12] (10) {10};

      \draw (11) edge[chartyellow, stealth-] (2) edge[chartyellow, stealth-] (5) edge[chartyellow, stealth-] (7) edge[chartyellow, stealth-] (9)
            (5) edge[chartmagenta, stealth-] (3)
            (9) edge[chartmagenta, stealth-] (6) edge[chartmagenta, stealth-] (8)
            (12) edge[chartyellow, stealth-] (10);
    \end{tikzpicture}
  \end{center}
  \caption{The \mt{s} computed for each commit in
    Figure~\ref{fig:repoDAG} showing the path that each commit takes to
    be merged into the master branch of the repository. This does not
    indicate how the events being merged are related. We retained the
    numerical order of the events, but the order can be arbitrary.}
  \label{fig:repoTree}
\end{figure}

@@

  \end{tikzpicture}
  \caption{The \mt applied to the inverted DAG to show the order that
    commits are created, as well as the merge sequence for the events in
  Figure~\ref{fig:repoEvents}. This requires a post-processing step
  involving the \mt produced by Algorithm~\ref{fig:alg} and the DAG.}
\label{fig:repoDAGTree}
\vspace{-3mm}
\end{figure}

@@

Computing the \mt from a DAG for any repository may not be possible;
however, certain features of the development process of Linux make it
feasible to compute the \mt for the Linux repository. First, the master
branch of Linux is maintained by Linus Torvalds, and only Linus has
write access to it. We have verified this assertion in previous
research~\cite{German2015}. We have developed a heuristic that is
presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
identifies the commits made directly to the master branch, whereafter it
recursively determines the shortest path, using the DAG, from each
commit to the master branch using the inverted DAG.\@

\begin{algorithm}
        \caption{Computing the \mt of Linux from the DAG}\label{fig:alg}
        \begin{algorithmic}[1]
                \Function{ComputeMergeTree}{DAG}: tree
                % \State {\# Compute the tree from the DAG of Linux repository.}
                % \State {\# Returns $Tree$, a graph containing every commit }
                % \State {\# in DAG with the path it followed to master.}
                \State $head \gets \textit{Head of master of git repository}$
                \State $master \gets \textit{traverse DAG from head using }$
                \State \quad\quad\quad\quad $\textit{first parent until reaching root}$
                \State $nodes(Tree) \gets nodes(DAG)$
                \State \Function{MergeAtMaster}{cid}
                \State {\# Returns $(depth, merge, next)$}
                \State {\# Helper function}
                \State {\# Compute the closest merge into master, }
                \State {\# setting the children on the way to master.}
                \If {\textit{cid in master}}
                \State \Return $(0, cid, \varnothing)$
                \EndIf
                \State {$d \gets \infty$}
                \State {\# Traverse the inverted DAG}
                \For{$c \in children(cid, DAG)$}
                \State $(d_c, merge_c, next_c) \gets MergeAtMaster(c)$
                \If {$IsMerge(c)$}
                \State $fp \gets FindFirstParent(c)$
                \If {$fp \neq cid$}
                \State $d_c \gets d_c + 1$
                \State $next_c \gets c$
                \EndIf
                \EndIf
                \State {\# Find the shortest path}
                \If {$d_c < d$}
                \State $(d, m, next) \gets (d_c, merge_c, next_c)$
                \ElsIf{ $d_c = d$ }

                \State {\# Use the time as a tie-breaker}
                \If {$ cTime(merge_c) < cTime(m) $}
                \State $(m, next) \gets (merge_c, next_c)$
                \EndIf
                \EndIf
                \EndFor
                \State {\# $c$ is the commit that follows $cid$}
                \State {\# in its way to master}
                \State add edge $(cid, next)$ to $Tree$
                \State \Return $(d, m, next)$
                \EndFunction

                \State {\# Compute the distance for each commit}
                \State {\# discarding result}
                \For{$c \in nodes(DAG)$}
                \State $MergeAtMaster(c)$
                \EndFor
                \State \Return $Tree$
                \EndFunction
        \end{algorithmic}
\end{algorithm}

The algorithm is broken into two phases. The first is determining which
repository events are on the master branch. This is done by traversing
the first parent from the master branch reference to the commit that has
no parents. The second phase is encompassed by the function
$MergeAtMaster$ which determines, for each commit, which merge the
commit is merged at, the depth (as variable $d$ in the algorithm), and
the next merge on the path to the master branch. The function
$MergeAtMaster$ has two parts, the first for determining the depth, from
the master branch, that the repository event is at. The second phase
determines the merge into the master branch, and the next merge on the
way to the master branch. The distance is by shortest path, staying as
close to the master branch as possible. If there is a tie between two
paths, the path that merges into the master branch sooner is taken.

To demonstrate the behaviour of the algorithm, we use a short example,
computing the merge at commit 5 in Figure~\ref{fig:repoEvents}.
$MergeAtMaster$, recurses along the children of the nodes it visits.
Eventually every child of every node along the path will be visited at
least once. Without loss of generality, suppose that the path recursed
along is from node 5 to 6, 9, 10, and finally 12. The depth for each,
except 12 (a merge into the master branch), is initialized to infinity,
the merge into master is blank, and the next merge is blank. Merges into
master trivially have a distance of 0 from the master branch, and it
merges itself into the master branch. The recursion at 12 returns the
triple $(0, \varnothing, 12)$ to the call from 10. 12 is a merge commit
and 10 is not the first parent, so the temporary depth, $d_c$, is
incremented to 1 and the temporary next merge, $next_c$, is changed to
12. 1 is less than infinity, so the depth is set to 1, the merge to 12,
and the next to 12. This returns the triple $(1, 12, 12)$ to the call
from 9. 9 is the first parent of 10, so no changes are made to the
temporary variables. The call to 9 recurses to the second child, 11. 11
is a merge into the master so it returns $(0, \varnothing, 11)$ to the
call from 9. 9 is not the first parent of 11, so the $d_c$ is
incremented to 1 and $next_c$ is changed to 11. The distances $d_c$ and
$d$ are the same, so to break the tie, we use the time. 12 was merged
after 11, so 11 replaces 12 as the merge into the master branch for 9,
as well as being the next merge. The call for 9 returns the triple $(1,
11, 11)$ to the call for 6. 6 is not the first parent of 9, so $d_c$ is
incremented and $next_c$ is changed to 9, as 9 merges 6. 2 is less than
infinity, so the $d$ is changed to 2, the merge to 11, and the next
merge to 9. The call to 6 returns the triple $(2, 11, 9)$ to the call
for 5. The call for 5 recurses on the second child of 5, calling on 7,
which calls 8, and then 9. 9 can continue, but if the implementation of
the algorithm uses memoization, the call to 9 can immediately return the
triple $(1, 11, 11)$ to the call for 8, and avoid an exponential
runtime. 8 is not the first parent of 8, so $d_c$ is incremented to 2
and $next_c$ is changed to 9. 2 is less than infinity, so $d$ is changed
to 2, $merge$ to 11, and $next$ to 9. The call to 8 returns the triple
$(2, 11, 9)$ to the call for 7, which recurses on the second child of 7,
9. 9 returns the triple $(1, 11, 11)$. 7 is the first parent of 9, so
the depth is not incremented. $d_c$ is less than $d$, so $d$ is changed
to 1, $m$ to 11, and $next$ to 11, returning $(1, 11, 11)$ to the call
for 5. $d_c$ is less than $d$, so $d$ is changed to 1, $m$ to 11, and
$next$ to 11. There are no other children, so the function halts.

@@

\begin{figure}[htbp]
        \centering
        \vspace{3mm}
        {\fontsize{7}{9}
        \begin{verbatim}
Merge: 8cbd84f fd8aa2c
Author: Linus Torvalds <torvalds@linux-foundation.org>
Date:   Tue Aug 10 15:38:19 2010 -0700

@@

* 'for-linus' of git://neil.brown.name/md: (24 commits)
md: clean up do_md_stop
[... edited for the sake of space]
md: split out md_rdev_init
md: be more careful setting MD_CHANGE_CLEAN
md/raid5: ensure we create a unique name for kmem_cache...
...
        \end{verbatim}}\vspace{-5mm}
        \caption{Example of how merges record a subset of commits being merged. The
                commit only shows the first 20 one-line summaries messages for the 24
                non-merge commits it merged. The ending ``\ldots'' is part of the log
                and represents that other commits were merged.}
        \label{fig:sampleMerge}
  \vspace{2mm}
\end{figure}

We used this information to evaluate the accuracy of the \mt model
extracted from the DAG\@. The method we followed started with the
extraction of the commit history up to July 20, 2016 form the Linux
repository. We computed the \mt of every commit until that date. Since
Linus Torvalds mostly does merging directly into the master branch, we
assumed that every merge by Linus is the root of a \mt, later detecting
merges that do not merge into the master branch and removing them from
the set of root merges. As described above, the log of a merge-commit
usually contains the number of commits in the merge the first 20
summaries of commits being merged. We extracted merges by Linus Torvalds
using the command \verb|git log --merges --author='Torvalds'| and
compared the number of commits stated in the log message with the number
of commits in the \mt. We also used the summaries
of the commits found in the merge (not necessarily all---see above) to
make sure those commits were in their corresponding \mt For example, for
the merge in Figure~\ref{fig:sampleMerge} we would expect that the \mt
rooted at \mycode{8cbd84f} contains 24 commits, and the one-line
summaries corresponds to commits in that \mt. We also inspected those
with differences to make sure they were true errors. The results can be
summarized as follows:
